\vspace{-2.0cm} \textbf{A self supervised method to detect nouns and numbers in a non-annotated signs language corpus}

### 2023-06-20 articles/2021/self-supervised-method-to-detect-nouns-and-numbers/article-tesis.tex
made updates to the introduction

\title{%
	\vspace{-2.0cm}
	\textbf{Talking with signs} \\	
	\Large \textbf{A self supervised method to detect nouns and numbers in a non-annotated signs language corpus}
}

@@

\begin{abstract}
Self supervised learning is a current topic that allows transferring learned latent variables on pre-text tasks into downstream tasks achieving notable increases in terms of model performance. In this work we present a self supervised approach for learning latent variables in a non annotated Peruvian Signs Language dataset using a SwAV based approach applied to video datasets. We present a novel signs detection model (nSDmV3) which takes advantage of a pre-trained self supervised features detection model. obtaining XX\% AUC To our knowledge it is the first work that uses SwAV as a baseline to detect signs in a non-annotated dataset. 
\\
\keywords{self supervised learning, SwAV, latent variables,Peruvian signs language}
\end{abstract}
\section{Introduction}\label{Introduction}
Starting a model architecture design by considering the usage of pre-trained models is always a good starting point,  even when using a ResNet50 network backbone pre-trained with ImageNet the resulting model will leverage a set of fine tuned layers decreasing hours and data volume required for training. There have been some efforts to build an annotated Peruvian Signs Language (PSL) dataset (Bejarano et al) with promising but still small results. In this paper we are using Self Supervised techniques to overcome the current low annotated PSL dataset availability problems towards to obtain improved results of a previous work that was focused on Supervised learning and classic movement and pattern recognition techniques \cite{Pereyra_2021_CVPR}.

Signs languages can be described from spatial-temporal features extracted from isolated glosses and a large set of continuous streams of contextually related frames.  Similar to other languages, context is extremely important to define what is semantically correct and to provide meaning to a set of interconnected glosses which relations can easily become extremely complex and can only be described from an also extremely large set of spatial and temporal features but most importantly latent variables intrinsic to the data.

This work is focused on non contrastive clustered methods in order to decrease the computation needed to compare feature by feature required for constrastive methods. Our method is based on Swapping Assignments between multiple Views (SwAV) \cite{caron2020unsupervised} as a backbone and extended it to design a set of model architectures that can learn latent variables from a set of non annotated PSL videos, training our model with the PSL dataset generated by the department of grammar at PUCP.

Sign Language corpus annotation is a tedious and complex task where a set of automatic and manual tasks need to be orchestrated together in order to produce a quality corpus with statistical value. In this work we solve the problem of lack of annotated PSL datasets availability by proposing a self supervised method where a features detection model can be trained to learn latent variables while solving a set of pre-task in our case finding compatible augmented video fragments across the dataset in order to solve a downstream nouns and numbers classification task.

Finally we extended model architectures that were previously created as part of \cite{Pereyra_2021_CVPR} to use our SwAV based pre-trained model as a backbone and to improve the AUC of PR metric in these models. This work is divided on Related Work \ref{related-work},  Experimentation \ref{experimentation},  Discussion \ref{discussion} and Conclusion \ref{conclusion}

\section{Related Work}\label{related-work}
\subsection{Self Supervised Learning}
Predicting the future or reconstructing the past is a way to learn and understanding the hidden patterns and distributions in data which is not necessarily annotated or pre-processed. Schmidhuber proposed an early self supervised method \cite{schmidhuber1990making} where two fully self supervised recurrent neural networks predicted future reinforcement based on external interactions and state sent to the next network cell is calculated on differences on successive predictions achieving a system capable to generate supervision signals. Unlike supervised tasks where predictions are influenced by target and observed variables, self supervised tasks are influenced in a greater extent by hidden or invisible variables also known as latent variables. By capturing those dependencies, a model can be used to answer questions about the values of unknown variables given the values of known variables. Energy-Based Models (EBMs) capture dependencies by associating a scalar energy (a measure of compatibility) to each configuration of the variables. Inference, i.e., making a prediction or decision, consists in setting the value of observed variables and finding values of the remaining variables that minimize the energy. Learning consists in finding an energy function that associates low energies to correct values of the remaining variables, and higher energies to incorrect values. A loss functional, minimized during learning, is used to measure the quality of the available energy functions. Within this common inference/learning framework, the wide choice of energy functions and loss functionals allows for the design of many types of statistical models, both probabilistic and non-probabilistic \cite{lecun2006tutorial}.

>>>>

\title{%
	\vspace{-2.0cm}
	\textbf{A self supervised method to detect nouns and numbers in a non-annotated signs language corpus} 
}

@@

\begin{abstract}
Self supervised learning is a current topic that allows transferring learned latent variables on pre-text tasks into downstream tasks achieving notable increases in terms of model performance. In this work we present a self supervised approach for learning latent variables in a non annotated Peruvian Signs Language dataset using a SwAV based approach applied to video datasets. We present a novel signs detection model (nSDmV3) which takes advantage of a pre-trained self supervised features detection model. obtaining XX\% AUC To our knowledge it is the first work that uses SwAV as a baseline to detect signs in a non-annotated dataset. 
\\
\keywords{self supervised learning, SwAV, latent variables,Peruvian signs language}
\end{abstract}
\section{Introduction}\label{Introduction}
In recent years, there has been a significant surge in interest surrounding inclusive software. The aim is to enhance accessibility and improve user experience. Incorporating inclusivity is essential to ensure seamless interaction between individuals with temporal or permanent physical limitations and computer systems. Deafness is a prevalent physical limitation, according to the World Health Organization (WHO), over 466 million people worldwide experience some level of hearing loss \cite{deafness_and_hearing_loss_2020} and approximately 8\% of the Peruvian population is affected by hearing loss \cite{disabilities_survey_2012}. Despite that there have been great advances on using AI for sign language recognition the lack of large annotated dataset and the complexity associated to learn intrinsic relations and dependencies between continuous sign languages constructions have posed challenges \cite{adaloglou2020comprehensive}. These challenges  have transformed the problem into a high dimensional problem that cannot be solved solely with supervision \cite{facebook-ssl}. Starting a model architecture design that leverages pre-trained models is always a good starting point even when using a ResNet50 pre-trained with ImageNet backbone, the resulting model will leverage a set of fine tuned layers decreasing time and data volume required for training. Previous knowledge and existing skill-sets are vital elements that will indeed decrease the teaching required to learn a new task. Yann Lecun have used an analogy where he compares the time and skills humans require to learn to drive a car with the massive amount of time and a data required to train autonomous cars emphasizing that previous knowledge and skillsets will greatly reduce training time as well as the data require to train a model \cite{facebook-ssl}.
In this work we use a non-annotated Peruvian Signs Language dataset created by the department of grammar at the PUCP, we will call it PSL dataset in the rest of this study. There have been some efforts to build an annotated Peruvian Signs Language (PSL) dataset (Bejarano et al) with promising results. However, there are challenges that are intrinsic to sign languages lexical and syntactical structures that require not only learning isolated glosses but continuous language constructions requiring to learn complex relationships and dependencies between actions that include several parts of the body including hands, arms, neck, head, eyes and eyebrows. 
Signs languages can be defined as a set of interrelated spatial-temporal elements composed from very large but still discrete set of isolated lexical glosses and most important from a continuous set of contextually related or interrelated streams of glosses that could be related to other streams of glosses recursively building complex continuous syntactical constructions.  
Continuous streams of glosses are high dimensional, which configures limitations when using probabilistic learning due to mathematical constraints while calculating probabilistic distributions using high dimensional data where integrals can become intractable\cite{kingma2022autoencoding}. In this paper we use the Sinkhorn distance which is an optimal transport distance we use to measure dissimilarity between continue streams of glosses probability distributions in an online and non-contrastive fashion, this to overcome the high dimensional curse by leveraging the geometry of the data.
As in every language context and common sense are extremely important to define what is semantically correct and to provide a meaning to a set of interconnected streams of glosses. Learning latent variables intrinsic to the dataset enables the development of common sense in the context of PSL dataset. This, in turn, significantly reduces the size of the annotated PSL dataset required for accurate classification of PSL elements such as nouns and numbers.
This paper employs self-supervision as a strategy to address the absence of a substantial dataset containing annotated PSL glosses and the nonexistence of a dataset comprising continuous streams of PSL language constructions during the training of our proposed novel signs language detection model nsvmv3. Annotating a Sign Language dataset is a laborious and intricate process that requires a combination of automatic and manual tasks to produce a high quality corpus with statistical value. While creating a large annotated PSL dataset is beyond the scope of this work, we propose a pretext task that leverage SwAV (Swapped assignment Views) \cite{caron2020unsupervised}. A self-supervised technique original designed for image data, we extended SwAV to work with the PSL dataset for training a spatial-temporal features model with the full PSL dataset. This approach allow us to learn latent variables from raw video files, enabling an understanding of the complex relations and dependencies between continuous PSL streams. This with the aim is to use our nsvmv3 model for a downstream task focused on successfully detecting and classifying signs utilizing small annotated PSL datasets for both training and inference purposes.
To to ensure scalability when using our method with large PSL datasets, this work uses non-contrastive or online clustering methods to reduce computational load when comparing features and generating prototypes as discussed earlier. Specially, we utilize the Sinkhorn distance as a non-constrastive and online method to compare prototypes probability distributions \cite{caron2020unsupervised}. 
The structure of this work comprises the following sections: Related Work (Section \ref{related-work}), Experimentation (Section \ref{experimentation}), Discussion (Section \ref{discussion}), and Conclusion (Section \ref{conclusion}).

\section{Related Work}\label{related-work}
\subsection{Self Supervised Learning}
Self Supervised Learning also called the dark matter of intelligence (Lecunn et all 2021) is a first step on human level AI since it is focused on learning a form of generalized common sense instead of learning limited features from a labeled datasets which leads to learn new skills without needing a great amount of teaching. Predicting the future or reconstructing the past is a way to learn and understanding the hidden patterns and distributions in data which is not necessarily annotated or pre-processed. Schmidhuber proposed an early self supervised method \cite{schmidhuber1990making} where two fully self supervised recurrent neural networks predicted future reinforcement based on external interactions and state sent to the next network cell is calculated on differences on successive predictions achieving a system capable to generate supervision signals. Unlike supervised tasks where predictions are influenced by target and observed variables, self supervised tasks are influenced in a greater extent by hidden or invisible variables also known as latent variables. By capturing those dependencies, a model can be used to answer questions about the values of unknown variables given the values of known variables. Energy-Based Models (EBMs) capture dependencies by associating a scalar energy (a measure of compatibility) to each configuration of the variables. Inference, i.e., making a prediction or decision, consists in setting the value of observed variables and finding values of the remaining variables that minimize the energy. Learning consists in finding an energy function that associates low energies to correct values of the remaining variables, and higher energies to incorrect values. A loss functional, minimized during learning, is used to measure the quality of the available energy functions. Within this common inference/learning framework, the wide choice of energy functions and loss functionals allows for the design of many types of statistical models, both probabilistic and non-probabilistic \cite{lecun2006tutorial}.
